---
title: T Wilcoxon Matched Pairs Signed Ranks Test
author: DocP
date: '2020-06-08'
slug: t-wilcoxon-matched-pairs-signed-ranks-test
categories: []
tags: []
---
```{r, echo = FALSE, message=FALSE}
  library(readr)
  BP <- read_csv("BloodPressure.csv")
  attach(BP)
  ```


Reference:Marin R Tutorial 4.5

Data set: BloodPressure

Read the data set in to R and attach it.
ASs always, it is a good idea to produce a graphic so we can better visualize our data. A box plot is appropriate in this case.

```{r}
boxplot(Before, After)
```

We may also want to see a summary of the data. I renamed it "BP" to save some typing.

```{r}
summary(BP)
```

The actual Wilcoxon matched pairs signed ranks test is as follows.

```{r}
wilcox.test(Before, After, mu = 0, alt = "two.sided", paired = T, conf.int = T, conf.level = 0.99, exact = F)
```

All are defaults except paired and exact. "paired = T" is necessary in this case so R knows to run the Wilcoxon T rather than the Rank Sums test. "exact = T" will generate warnings due to the ties.

The small p value suggests that whatever manipulation we performed between the Before and After readings were taken must have been successful.

