---
title: M Checking Regression Assumptions
author: Doc P
date: '2020-06-10'
slug: m-checking-regression-assumptions
categories: []
tags: []
---
 ```{r, echo = FALSE, message=FALSE}
  library(readr)
  LCD <- read_csv("LungCapData.csv")
  attach(LCD)
  ```

Reference: Marin Video 5.2

WHile we will not be doing much with this information, there are several easy tests we can perform to deteremine whether a linear regression is a proper approach for eplaining a relationship.

First, Import and attach the LungCapData set.

Next, run the code we used in the last Crib Sheet to generate the regression analysis.

```{r}
plot(Age, LungCap)
mod<-lm(LungCap ~Age) 
summary(mod) 
abline(mod)
```


At this point "plot(mod)" will produce the diagnostic plots, hit return to cycle through the plots.

```{r}
plot(mod)
```

1st plot is a residual plot - the red line should be straight and residuals randomly distributed.

2nd plot is a Q-Q plot (Quantile - Quantile) and will be a diagonal straight line if residuals are normally distributed.

3rd and 4th identify other problems with which we are not concerned.

Non constant variance will show up in the plots with a “megaphone” shape, while non-linearity will show up with a curved line on the diagnostic plots.  Our plots look good, suggesting that a linear regression analysis is a reasonable approach for these data.